---
title: 'Homework #4'
author: Gabrielle Scibelli 
output: 
  html_document:
    df_print: paged 
    toc: true 
    toc_depth: 2
    toc_float: true 
    number_sections: true
    theme: united 
    highlight: espresso 
---


# gganimate 
The following link is for the data that is used in this Section. 

```{r setup,echo=FALSE}
library(ggplot2)
library(gganimate
        )
imf_growth <- read.csv(url('https://bcdanl.github.io/data/imf_growth_quarterly.csv'))
summary(imf_growth)
imf <- ggplot(data = imf_growth, 
              mapping = aes(x = quarter, y = reorder(country, gy),
                            fill = gy))
  
imf + geom_col()
  
```

- geom_tile() and geom_text() 


## ggiraph 
The following link is for the data that is used to generate the figure in this Section. 
```{r cars}
summary(cars)
```

The effect of the COVID-19 Pandemic on economic growth is shown in the animation above. 

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
